% !TEX root = ../TAMU_Thesis_Main.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                           ACKNOWLEDGMENTS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{acknowledgements}
This dissertation is the result of support from dozens of dear friends, mentors, and colleagues since I started doing research in 2017.

First, I am deeply thankful to Dr.~Robert Hetland for his mentorship since taking me on as an REU student in 2018. He introduced me to ocean modeling and ignited my passion for oceanography. He gave me the freedom to pursue my own research interests and supported me through every setback and discovery. He fostered my growth as a scientist and I would not be the man I am today without his support. 

Second, I am grateful to my committee, Drs.~Henry Potter, Spencer Jones, and Scott Socolofsky for their critical guidance that helped shape my dissertation. My work sits at the intersection of numerics and physics and their support was instrumental in communicating these concepts to a general oceanographic audience. I am especially grateful to Spencer Jones for our discussions over the years about numerical mixing, my career, and the tribulations that come with living in College Station without a vehicle. He provided me the opportunity to mentor students and helped foster a sense of community when I needed it most. 

Third, I am grateful to the Submesoscales Under Near-Resonant Inertial Shear Experiment (SUNRISE, NSF Grant OCE-1851470) team for giving me the opportunity to collaborate with world class scientists. Our student meeting in Bend, OR during 2022 was a highlight of my graduate experience, as it was my first time visiting the west coast. 

Fourth, I am grateful to the Office of Science Graduate Student Research (SCGSR) Program for giving me the opportunity to work at Los Alamos National Laboratory with the COSIM group. They provided funding for the work presented in Chapter 4. This was a life changing experience because Los Alamos is the most beautiful place I have lived and allowed me to finally explore my outdoor hobbies and interests. In addition, I worked on a new research project that demanded the very best of me and was supported by half a dozen world class researchers and wonderful humans. I am thankful to Drs.~Mark Petersen, Katherine Smith, and LeAnn Conlon for their mentorship throughout the project, which helped me accomplish a ridiculous amount of work in just four months. In particular, I am thankful to Mark Petersen for being so kind to me after I first moved to Los Alamos and for pushing me to become a social runner. I thoroughly enjoyed our runs through the streets and trails of Los Alamos. 

Last, I am thankful to Kaila Uyeda for giving me the chance to be a mentor. She gave me a reason to be better and supported me as I scrambled to write my second chapter and struggled to figure out what comes after graduate school. 
\end{acknowledgements}