% !TEX root = ../TAMU_Thesis_Main.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%             CONTRIBUTORS AND FUNDING SOURCES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{contributors}
\subsection*{Contributors}
This work was supported by a dissertation committee consisting of Drs.~Henry Potter (Chair), Robert Hetland (Co-chair), and Spencer Jones of the Department of Oceanography, and Dr.~Scott Socolofsky of the Department of Civil Engineering.

For the work published in Chapter 2, the student, Dr.~Hetland and Dr.~Lixin Qu of Shanghai Jiao Tong University designed and performed the research. Dr.~Daijiro Kobashi of Fugro N.V. designed the numerical simulations. The student analyzed the numerical simulations. All wrote the paper. 

For the manuscript under review for publication in Chapter 3, the student and Dr.~Hetland designed and performed the research. Dr.~Daijiro Kobashi designed the realistic numerical simulations referenced in the manuscript. The student developed the idealized numerical simulations used in the manuscript. Dr.~Spencer Jones of the Department of Oceanography aided in analysis/interpretation of the numerical simulations. All wrote the paper. 

Project design and numerical simulation conceptualization, development, and analysis for Chapter 4 were contributed to by Drs.~Mark Petersen, Katherine Smith, LeAnn Conlon, Matthew Maltrud, and Darren Engwirda of Los Alamos National Laboratory and Dr. Hetland.  Drs.~Petersen, Smith, and Hetland contributed to the text in Chapter 4. 

The student independently completed all other work for the dissertation.

\subsection*{Funding Sources}
Dylan Schlichting was fully supported by National Science Foundation Grant OCE-1851470 for the first four years of the Ph.D.~program. For the final eight months of the Ph.D.~program, the Office of Science Graduate Student Research (SCGSR) Program provided salary and relocation support at Los Alamos National Laboratory. For the final 8 months of the Ph.D.~program, the same NSF grant provided benefits, tuition, and fees support. 
\end{contributors}